
---
title: "Statistical Analysis with R: Midterm Presentation"
author: <font size="6"> Matt Melnychuk </font>
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      
class: inverse, middle, center
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)

#xaringanExtra::use_fit_screen()
```
class: inverse, middle, center
# Graphs
---
# Questions Examined:

--

- Do Hospitals Specialize?

--

- What makes patients stay longer?

--

- What impacts birth weight?

--

- Which hospitals cost the most?


---
# Do Hospitals Specialize?

- Age?

- Sex?

- Type of Service/Admittance?

---
```{r, echo=FALSE, message=FALSE}

library(tidyverse)
df <- read_csv("C:/Users/student/Desktop/R 421/MidtermClean.csv")
df2 <-  select(df, -c('payfix','asource','preopday','bwght','ecodub92','pt_state','diag_adm','er_mode','obs_hour','nicu_day'))
df2$hospital <- case_when(
  df2$provider == '7201' ~ 'Newport',
  df2$provider == '7202' ~ 'St. Joseph',
  df2$provider == '7203' ~ 'Memorial',
  df2$provider == '7204' ~ 'Miriam',
  df2$provider == '7205' ~ 'Rhode Island Hospital',
  df2$provider == '7206' ~ 'Roger Williams',
  df2$provider == '7209' ~ 'South County',
  df2$provider == '7210' ~ 'Kent County',
  df2$provider == '7211' ~ 'Westerly',
  df2$provider == '7212' ~ 'Rehab of RI',
  df2$provider == '7213' ~ 'Landmark Medical Center',
  df2$provider == '7214' ~ 'Women and Infants',
  df2$provider == '7215' ~ 'Bradley',
  df2$provider == '7216' ~ 'Butler',
  TRUE ~ '')

df2$RE2 <- case_when(
  df2$raceethn == '1' ~ 'White',
  df2$raceethn == '2' ~ 'Black',
  df2$raceethn == '3' ~ 'Asian',
  df2$raceethn == '4' ~ 'American Indian',
  df2$raceethn == '5' ~ 'Native Hawaiian',
  df2$raceethn == '6' ~ 'Other',
  df2$raceethn == '7' ~ 'Hispanic',
  df2$raceethn == '9' ~ 'Unknown',
  TRUE ~ '')

df2$SEX2 <- case_when(
  df2$sex == '1' ~ 'Male',
  df2$sex == '2' ~ 'Female',
  TRUE ~ 'Unknown')

df2$Camp2 <- case_when(
  df2$campus == '1' ~ 'St. Joseph - Fatima',
  df2$campus == '2' ~ 'St. Joseph - Specialty',
  df2$campus == '3' ~ 'RI Adult',
  df2$campus == '4' ~ 'RI Hasbro',
  df2$campus == '0' ~ 'Only Campus',
  TRUE ~ '')

df2$Admit2 <- case_when(
  df2$admtype == '1' ~ 'Emergency',
  df2$admtype == '2' ~ 'Urgent',
  df2$admtype == '3' ~ 'Electric',
  df2$admtype == '4' ~ 'Newborn',
  df2$admtype == '5' ~ 'Trauma',
  df2$admtype == '9' ~ 'Unknown',
  TRUE ~ '')
```

# Age by Hospital

```{r, echo=FALSE, out.width = "50%"}
library(tidyverse)
df2 %>% ggplot()+ geom_boxplot(mapping = aes(y = age, color = hospital))+ labs(title = 'Age by Hospital', x='', y='Age', fill='Hospital', caption = 'Bradley and W&I much younger.')
```

- Bradley focuses exclusively on younger patients.

- Women and Infants also has lots of young patients and newborns.

- Hasbro (`Campus 4`) also focuses mainly on treating children.

---
# Sex by Hospital

```{r, echo=FALSE, out.width = "50%"}
library(tidyverse)
df2 %>% ggplot()+geom_bar(mapping=aes(x=hospital, fill=SEX2), 
           position = 'fill')+
    labs(title = 'Sex by Hospital', x='', y='Proportion', fill='Sex', caption = 'W&I 75% female')
```
 
- Women and Infants focuses mainly on women

- All other hospitals are split mostly 50/50

---
# Admittance by Hospital

```{r, echo=FALSE, out.width = "50%"}
df2 %>% ggplot()+geom_bar(mapping=aes(x=hospital, fill=Admit2),position = 'fill')+
    labs(title = 'Admittance Reason by Hospital', x='', y='Proportion', fill='Hospital', caption = 'Bradley and Butler more urgent than emergency')
```

- Bradley and Butler have lots of urgent

- Maybe because Bradley and Butler are psychiatric hospitals

- Women and Infants has lots of newborns

---
# ICU Length by Hospital

```{r, echo=FALSE, out.width = "50%"}
library(ggplot2)
df2$icu <-as.numeric(df2$icu)
df2$icu <- as.numeric(df2$icu/100)
df2 %>% ggplot()+
  geom_boxplot(mapping = aes(y = icu, color = hospital)) + labs(title = 'ICU stays by hospital', x='', y='Days', fill='Hospital', caption = 'Kent County has long ICU stays.')
```

- Longest intensive care stays in Kent County hospital

- Maybe they are better equipped for these specific needs

---
# Why do patients stay for a long time?

- Age? 

- Race / ethnicity?

- Impact of Hospital?

---
# Length of Stay by Age & Race / Ethnicity?

```{r, echo=FALSE, out.width = "50%"}
df2  %>% ggplot()+
  geom_point(mapping = aes(x = age, y = los, color = RE2)) + labs(title = 'Length of Stay by race and age', x='Age', y='Length of Stay', caption = 'No patterns immediately noticeable')
```

- No obvious patterns to race

- Most lengths shorter than 75 days, a few longer term stays

- Older patients are more short-term, no other obvious patterns


---
# Length of Stay by Hospital

```{r, echo=FALSE, out.width = "50%"}
df2  %>% ggplot()+
  geom_boxplot(mapping = aes(y = log(los), color = hospital)) + labs(title = 'Length of Stay by Hospital', y='Length of Stay', fill='Hospital', caption = 'Bradley and Butler slightly longer stays.')
```

- Used log(length of stay) to help with visuals

- Bradley and Butler longer stays (psych hospitals again)

- South County & Women and Infants are the shortest lengths

---
# What impacts birth weight?

- Hospital of birth?

- Month?

- Race / ethnicity?

- Sex?

---
# Weight of baby by Hospital and Month
```{r, echo=FALSE, out.width = "50%"}
df2$b_wt <- as.numeric(df2$b_wt)
df2 %>%  filter(b_wt>0) %>%  filter(Admit2 =='Newborn')%>% ggplot() + geom_smooth(mapping=aes(x=moa, y=b_wt)) + facet_wrap(vars(hospital)) + labs(title = 'Newborn Weight by Month by Hospital', x='Month', y='Weight (g)', caption = 'Smaller babies at W&I.')
```

- No monthly pattern to weights

- Babies born at Women and Infants are smaller

- Maybe Women and Infants has more resources for premature babies

---
# Birthweight by sex and race
```{r, echo=FALSE, out.width = "50%"}
df2 %>%  filter(b_wt>0) %>% filter(SEX2 !='Unknown') %>% ggplot()+
  geom_boxplot(mapping = aes(x=SEX2, y = b_wt, color = RE2)) + labs(title = 'Birth Weight by Race', x='Sex', y='Weight (g)', fill='Race/Ethnicity', caption = 'No significant difference in weight by race or sex.')
```

- No significant differences by race or sex

---
# Why are hospital bills so expensive?

- Distribution of total bill?

- Age?

- Reason for admittance?

- Hospital?

---
# Distribution of Total Cost
```{r, echo=FALSE, out.width = "50%"}
df2 %>% filter (tot <=200000) %>% ggplot()+
  geom_histogram(binwidth = 1000, mapping = aes(x = tot)) + labs(title = 'Histogram of Hospital Bills', x='Cost', y='Count', caption = 'Mostly lower bills, some right skew.')
```

- Right skewed (some very large bills)

- Most bills under ~$35,000

---
# Average Cost by Hospital by Age
```{r, echo=FALSE, out.width = "50%"}
df2 %>% ggplot()+geom_smooth(mapping = aes(x = age, y = tot, color = hospital)) + labs(title = 'Average Cost by Hospital by Age', x='Age', y='Cost', caption = 'Bradley is the most expensive.')
```

- Bradley most expensive, RI Hospital second most

- Middle-aged patients typically pay more (based on parabolic curve)

- Butler is actually cheaper for middle-aged patients

---
# Average Cost by Admittance Type
```{r, echo=FALSE, out.width = "50%"}
df2  %>% filter(Admit2 != '') %>% filter(Admit2 != 'Unknown') %>% ggplot()+ geom_boxplot(mapping = aes(y = log(tot), color = Admit2)) + labs(title = 'Cost by Admittance Type', y='Log of Cost', fill='Admittance Type', caption = 'Trauma costs more on average.')
```

- Trauma costs more

- Newborns cost less
---
# Average Cost per Day by Admittance Type
```{r, echo=FALSE, out.width = "50%"}
df2$costperday <- df2$tot/df2$los
df2  %>% filter(Admit2 != '') %>% filter(Admit2 != 'Unknown') %>% ggplot()+ geom_boxplot(mapping = aes(y = log(costperday), color = Admit2)) + labs(title = 'Cost by Admittance Type', y='Log of Cost', fill='Admittance Type', caption = 'Newborns much cheaper per day.')
```

- Newborn visits cost much less than average on a per day basis

- Trauma costs now more in-line with other reasons

- Maybe trauma care itself isn't too expensive, just long time in hospital

---
# Average Cost Per Day by Hospital
```{r, echo=FALSE, out.width = "50%"}
df2$lncpd <- log(df2$costperday)
df2 %>% filter(costperday>0) %>%ggplot()+
  geom_density(mapping = aes(x = lncpd, color = hospital)) + labs(title = 'Natural Log of Cost per day by Hospital', x='ln(costperday)', y='Density', fill='Hospital', caption = 'Butler and Bradley average less per day.')
```

- Used log transformation to handle right skew

- Bradley and Butler cost much less per day

---
# Average Cost Per Day by Hospital

```{r, echo=FALSE, out.width = "50%"}
df2$lncpd <- log(df2$costperday)
df2 %>% filter(costperday>0) %>% filter(hospital !='Bradley') %>% filter(hospital !='Butler') %>%ggplot()+
  geom_density(mapping = aes(x = lncpd, color = hospital)) + labs(title = 'Natural Log of Cost per day by Hospital', x='ln(costperday)', y='Density', fill='Hospital', caption = 'St. J and W&I are slightly cheaper per day.')
```

- St. Josephs and Women and Infants slightly cheaper per day

- This may be caused by many W&I patients having the reason as newborn

- Rest of hospitals have similar cost distributions

---
# Revealing Line Graph
```{r, echo=FALSE, out.width = "50%"}
library(ggplot2)
library(gganimate)
df2$moa <- as.numeric(df2$moa)
df2 %>%  group_by(hospital, moa) %>% summarise(TotalCost = sum(tot)) %>% 
  ggplot(aes(y=TotalCost,
            x=moa, color=hospital))+ 
  geom_line()+
  transition_reveal(moa) + labs(x='Month', 
       y = 'Revenue from All Patients', 
       color='Provider', 
       title = 'Revenue by Provider by Month of Admittance', 
       caption = 'RI Hospital receives the most $')
```

- Rhode Island Hospital consistently receives the most revenue

- No apparent pattern to cash flow (no dramatic seasonal changes)

---
# Animated Bar Race

```{r, echo=FALSE, out.width = "50%"}
d1 <- df2 %>% filter(Admit2 == 'Newborn',hospital!='Women and Infants') %>% group_by(hospital, moa) %>% summarise(births = n())
d2 <- d1 %>% group_by(moa) %>% mutate(rank=rank(-births)) %>% ungroup()
a1 <- d2 %>% ggplot(aes(x=rank, y=births, group=hospital, fill=hospital, label=hospital)) + geom_col()+
    geom_text(aes(y = births, label = hospital), hjust = 1.4)+ 
    coord_flip(clip = "off", expand = FALSE) +scale_x_reverse()+
    labs(title = 'Month {closest_state}', x='', y='Total Number of Births', fill='Hospital', caption = 'Births by Provider (exclduing Women and Infants)')+
    theme(plot.title = element_text(hjust = 1, size = 22),
          axis.ticks.y = element_blank(),
          axis.text.y  = element_blank()) + 
    transition_states(moa)+
    ease_aes("cubic-in-out")
library(gifski)
animate(a1, renderer = gifski_renderer())

```

- Ran plot without W& I; always have most births (~750 per month)

- Kent County usually more than South County (except Feb and March)

- Maybe babies born here if mothers are not located close to Providence

---
class: inverse, middle, center
# Predicitive Modeling Section

---
# Model for High Cost

```{r, echo=FALSE, out.width = "42%"}
df2$target <- case_when(
  df2$tot < 21854 ~ 'low',
  df2$tot >= 21854 ~'high',
  TRUE ~ 'high')
library(tidyverse)
df3 <- df2 %>% filter(raceethn!='') %>% filter(admtype!='')
df3 <- select(df3, c("age","sex","raceethn","provider", "moa","mod","campus", 'los','target','admtype'))

#categorical variables
df3$sex <- as.factor(df3$sex)
df3$raceethn <- as.factor(df3$raceethn)
df3$provider <- as.factor(df3$provider)
df3$moa <- as.factor(df3$moa)
df3$mod <- as.factor(df3$mod)
df3$campus <- as.factor(df3$campus)
df3$target <- as.factor(df3$target)
df3$admtype <-as.factor(df3$admtype)

#continuous variables
df3$los <- as.numeric(df3$los)
df3$age <- as.numeric(df3$age)

library(caret)
set.seed(2020)
splitIndex <- createDataPartition(df3$target, p = .90, 
                                  list = FALSE)
df3_train <- df3[ splitIndex,]
df3_test <- df3[-splitIndex,]

set.seed(2020)
df3small <- sample_n(df3,13000)
library(caret)
set.seed(2020)
splitIndex <- createDataPartition(df3small$target, p = .90, 
                                  list = FALSE)
df3_train <- df3small[ splitIndex,]
df3_test <- df3small[-splitIndex,]

trControl = trainControl(method = "cv",
                         number = 5)

model1 <- train(target~., data=df3_train, 
                                method = "rocc", 
                                trControl = trControl)
model2 <- train(target~., data=df3_train, 
                    method = "pam", 
                                trControl = trControl)

tuneGrid = expand.grid(mtry = 2:5,
                       splitrule = c('gini', 'extratrees'),
                       min.node.size = c(5:10))
set.seed(2020)
model3 <- train(target~., data=df3_train, 
                    method = "ranger", 
                    trControl = trControl,
                    tuneGrid = tuneGrid)

results <- resamples(list('ROC-Based Classifier' = model1,
                          'Nearest Shrunken Centroids'= model2,
                          'Ranger Forest'=model3))
bwplot(results)

```

- Predicting High Cost in 3 methods

- Ranger method looks best

- Need to further examine which hyper parameters to use

---
# Final Model Decision and Accuracy

```{r, echo=FALSE, out.width = "50%"}
set.seed(2020)
model3 <- train(target~., data=df3_train, 
                    method = "ranger", 
                    trControl = trControl,
                    tuneGrid = tuneGrid)
plot(model3)
```

- Gini splits, 5 variables to choose, min node = 9

- Accuracy = 84.06%

---
# My Model of Choice

- Similar idea, binary test

- Want to examine what determines if length of stay is greater than median

- Median is 3 days

- Check if these models agree with graphs 

---
# Decision Tree
```{r, echo=FALSE, out.width = "50%"}
df3$targetB <- case_when(
  df3$los < 3 ~ 'low',
  df3$los >= 3 ~'high',
  TRUE ~ 'high')

df3$targetB <- as.factor(df3$targetB)
df3B <- select(df3,-c('los','target'))

#back to using full data set
library(caret)
set.seed(2020)
splitIndex <- createDataPartition(df3B$targetB, p = .90, 
                                  list = FALSE)
df3_train <- df3B[ splitIndex,]
df3_test <- df3B[-splitIndex,]

library(rpart)
set.seed(2020)
tree_model <- rpart(targetB ~ ., data = df3_train,
                 control = rpart.control(maxdepth = 4))

# Plot the results
library(rattle)
fancyRpartPlot(tree_model)
```

- First split is admittance type: emergency, unknown, trauma, and other lead to long stays

- Next split is hospital treating the patient

---
# Variable Importance
```{r, echo=FALSE, out.width = "42%"}
barplot(tree_model$variable.importance)
```

- Most important variable is reason for admittance (1st split)

- Next most important are provider (and campus) and age, this agrees with charts earlier

- Race plays a very small role, rest of variables have no impact

---
# Hyperparameter Selection in Ranger

```{r, echo=FALSE, out.width = "50%"}
set.seed(2020)
df3Bsmall <- sample_n(df3B,13000)
library(caret)
set.seed(2020)
splitIndex <- createDataPartition(df3Bsmall$targetB, p = .90, 
                                  list = FALSE)
df3_train <- df3Bsmall[ splitIndex,]
df3_test <- df3Bsmall[-splitIndex,]

trControl = trainControl(method = "cv",
                         number = 5)

model1 <- train(targetB~., data=df3_train, 
                                method = "rocc", 
                                trControl = trControl)

model2 <- train(targetB~., data=df3_train, 
                    method = "pam", 
                                trControl = trControl)

tuneGrid = expand.grid(mtry = 2:5,
                       splitrule = c('gini', 'extratrees'),
                       min.node.size = c(5:10))
set.seed(2020)
model3 <- train(targetB~., data=df3_train, 
                    method = "ranger", 
                    trControl = trControl,
                    tuneGrid = tuneGrid)
plot(model3)

```

- Gini split better than extra trees again (but not by as much as before)

- 5 variable predictors at each step, minimum node size = 8

---
# Model Comparison
```{r, echo=FALSE, out.width = "50%"}
results <- resamples(list('ROC-Based Classifier' = model1,
                          'Nearest Shrunken Centroids'= model2,
                          'Ranger Forest'=model3))
bwplot(results)
```

- Random forest (`ranger`) is the best model again

- Testing accuracy of 65.51%

---
class: inverse, middle, center
# Errors/ Issues

---
# Errors / Issues

--

- interpretability

    - used Google doc to connect variable codes to name (provider, sex, reason for admittance)
    
--

- large data set

    - used a random sample for some of the modeling as needed

--

- skewness

    - used log(length of stay) and log(total cost) on some graphs

--

- When creating the PowerPoint, I could not properly download xaringanExtra to use fit to screen

    - Sized images to 50% using alternative method


---
class: inverse, middle, center
# Thank You!
### Any Questions?
##### November 11, 2021

