
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 6: Visualization - Aesthetic Mapping"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](fa2021_assignment6.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

### Questions on Your Selected Data

Choose a dataset and produce the following types of plots and comment on each plot. Plots should be meaningful. If you use the data we used in class, make sure the plots are not the same as the ones in the slides. 

-------
```{r}
library(tidyverse)
library(ggplot2)
df6 <-read_csv("C:/Users/student/Desktop/Data Mining 460/ncaaVB.csv")
```


1. For one continuous variable:

- Density
```{r}
df6 %>% ggplot()+
  geom_density(mapping = aes(x = AttackErrors))
#Attack errors appear to be normally disctributed for NCAA teams. The mean is around 5.4 errors per set.
```

- Histogram 
```{r}
df6 %>% ggplot()+
  geom_histogram(mapping = aes(x = Digs))
# In general, the number of digs per set is normally distributed for NCAA teams. The most frequent bin is around 15 digs per set (30+ teams fit in this bin).  All values fall between 10 and 20 digs per set. 
```

  
- Boxplot
```{r}
df6 %>% ggplot()+
  geom_boxplot(mapping = aes(y = Blocks))
# The median number of blocks is around 2 per set.  Q1 is about 1.75, and Q3 is about 2.25. There is one team with over 3 blocks per set and one team with less than 1 block per set.
```

-------

2. For one categorical variable

- Barplot
```{r}
df6 %>% ggplot()+
  geom_bar(mapping = aes(x = CoachLength))
#Not many coaches last for over three decades. The majority of schools have coaches who have been coaching for some years (between 3-8 years). Around 45 schools have new coaches with 0-2 years of coaching experience.
```


-------

3. For one two continuous variables

- Scatter Plot
```{r}
df6  %>% ggplot()+
  geom_point(mapping = aes(x = HitPercent, y = WinPercent))
#This scatterplot shows that there is a high correlation between a team's hitting percentage and a team's winning percentage. This makes sense because team's that are more effecient when attacking the ball tend to win more points which leads to more victories in matches.
```

- Line Plot
```{r}
df6 %>% filter(Conference=='NEC'|Conference=='Ivy League'|Conference=='SEC') %>% ggplot()+
  geom_line(mapping = aes(x = Aces, y = WinPercent, color = Conference))
# For NEC teams, it appears that having more Aces leads to a better records of wins and losses. The SEC line plot is a bit random seeming, and the Ivy League line plot shows that the team with the most Aces had the lowest winning percentage, which does not make sense.
```

- Smooth Plot
```{r}
df6 %>% ggplot()+
  geom_smooth(mapping = aes(x = Digs, y = ReceptionErros, color = Region))
#Throughout the country, teams that average more digs per set have less reception errors. This makes sense because a good defensive team will have lots of digs and few reception errors. The steepest curve occurs in the East region.
```

-------

4. For one continuous + one categorical variables

- Density
```{r}
df6 %>%  filter(Conference=='NEC'|Conference=='Ivy League'|Conference=='SEC') %>% ggplot()+
  geom_density(mapping = aes(x = Attacks, color = Conference))
#NEC Schools are more likely to have a higher number of attacks per set.  The distribution for the Ivy League and SEC look somewhat normally distributed, but Ivy League has a slight skew to the left, and the SEC has two peaks around the center.
```


- BoxPlot
```{r}
df6 %>% filter(Conference=='NEC'|Conference=='Ivy League'|Conference=='SEC') %>% ggplot()+
  geom_boxplot(mapping = aes(y = Sets, color = Conference))
#Ivy League schools played less sets during the 2019 season than NEC or SEC schools, possibly because the conference has fewer teams.  SEC play around 114 sets on average, while NEC teams play around 107 sets. At least one NEC school played in very few sets. This is Merrimack: 2019 was their first year in Division I so they did not play many additional non-conference matches.
```


-------

5. For two categorical variables: barplot
```{r}
df6 %>% filter(Conference=='NEC'|Conference=='Ivy League'|Conference=='SEC') %>% ggplot()+
  geom_bar(mapping = aes(x = CoachLength, fill = Conference))
#No Ivy League coaches have been at their current role for over 3 decades.  Asides from that, it appears that the distribution of coaching tenure is very similar regardless of the conference in which the college/university competes.
```


-------

### Questions on Covid Vaccination Data

In this section of questions, we will use  the covid vaccination data for the US. The data and data dictionary can be found at the below link:

- [Data](https://raw.githubusercontent.com/govex/COVID-19/master/data_tables/vaccine_data/us_data/time_series/vaccine_data_us_timeline.csv)

- [Data Dictionary](https://github.com/govex/COVID-19/blob/master/data_tables/vaccine_data/us_data/data_dictionary.csv)


```{r}
library(tidyverse)
df <- read_csv('https://raw.githubusercontent.com/govex/COVID-19/master/data_tables/vaccine_data/us_data/time_series/vaccine_data_us_timeline.csv')
```

1. `Geom_point` makes scatter plots, which plot two numeric (or Date) variables. A third categorical variable is often brought in for color. For example, we can compare different vaccine types in Rhode Island by

```{r}
df %>% filter(Province_State=='Rhode Island') %>% ggplot()+geom_point(mapping = aes(x = Date, y = Doses_admin, color = Vaccine_Type))
```

Do the follows. 

  - Use `geom_point` to compare the the Moderna and Pfizer vaccine administered in California
```{r}
df %>% filter(Province_State=='California') %>% filter(Vaccine_Type==c('Moderna','Pfizer')) %>% ggplot()+geom_point(mapping = aes(x = Date, y = Doses_admin, color = Vaccine_Type))
#Many more Pfizer administered
```
  
  - Use `geom_point` to compare the total vaccine administered in Rhode Island and Massachusetts
```{r}
df %>% filter(Province_State==c('Rhode Island','Massachusetts'),Vaccine_Type=='All') %>% ggplot()+geom_point(mapping = aes(x = Date, y = Doses_admin, color = Province_State))
```
  
  - Change `geom_point` to `geom_line` in the above codes to have the corresponding line plots. 
```{r}
df %>% filter(Province_State=='California') %>% filter(Vaccine_Type==c('Moderna','Pfizer')) %>% ggplot()+geom_line(mapping = aes(x = Date, y = Doses_admin, color = Vaccine_Type))
#Many more Pfizer administered

df %>% filter(Province_State==c('Rhode Island','Massachusetts'),Vaccine_Type=='All') %>% ggplot()+geom_line(mapping = aes(x = Date, y = Doses_admin, color = Province_State))
```
  

2. Sometime we need to do some calculations on the data before plotting. For example, to compare the popularity of two vaccine `Moderna` and `Pfizer` by date, we can use. 

```{r}
# Create a dataset recording these two vaccine administered by date
d1 <- df %>% filter(Vaccine_Type==c('Moderna','Pfizer')) %>% group_by(Date, Vaccine_Type) %>% summarise(vac_admin=mean(Doses_admin, na.rm=TRUE))

# Plot from the newly create data
d1 %>% ggplot()+geom_point(mapping=aes(x=Date, y=vac_admin, color=Vaccine_Type))
```

Do the follows. 

- Use `geom_point` to compare the popularity of all vaccines by date.
```{r}
df  %>% filter(Vaccine_Type ==c('Janssen','Moderna','Pfizer','Unknown','Unassigned')) %>% group_by(Date, Vaccine_Type) %>% summarise(DOSES = mean(Doses_admin)) %>% ggplot()+geom_point(mapping = aes(x = Date, y = DOSES, color = Vaccine_Type))
```

- Use `geom_line` to compare the popularity of all vaccines in New England by date.
```{r}
df  %>% filter(Vaccine_Type ==c('Janssen','Moderna','Pfizer','Unknown','Unassigned')) %>% filter(Province_State==c('Rhode Island','Maine','Connecticut','New Hampshire','Massachusetts','Vermont')) %>%  group_by(Date, Vaccine_Type) %>% summarise(DOSES = mean(Doses_admin)) %>% ggplot()+geom_line(mapping = aes(x = Date, y = DOSES, color = Vaccine_Type))
#Gaps in days make the graph jump down to 0 then back up to a more relevant value
```

- Use `geom_line` to compare the popularity of all vaccines in New England by month. 
```{r}
library(lubridate)
df  %>% filter(Vaccine_Type ==c('Janssen','Moderna','Pfizer','Unknown','Unassigned')) %>% filter(Province_State==c('Rhode Island','Maine', 'Connecticut','New Hampshire','Massachusetts','Vermont')) %>%  group_by(Date, Vaccine_Type) %>% summarise(DOSES = mean(Doses_admin)) %>% ggplot()+geom_line(mapping = aes(x = month(Date), y = DOSES, color = Vaccine_Type))
#Peaks at whole integers for month values
```

- Use `geom_line` to compare states in New England in term of the total vaccine administered by date.
```{r}
df  %>% filter(Vaccine_Type ==c('All')) %>% filter(Province_State==c('Rhode Island','Maine','Connecticut','New Hampshire','Massachusetts','Vermont')) %>%  group_by(Date, Vaccine_Type, Province_State) %>% summarise(DOSES = mean(Doses_admin)) %>% ggplot()+geom_line(mapping = aes(x = month(Date), y = DOSES, color = Province_State))
#Massachusetts has most vaccines - makes sense because also most people
```
