
---
output: # rmarkdown::github_document
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 3.  Working with Data Frame. Base R Style"
---

***Note*:** *This assignment practices working with Data Frame using Base R.*

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](fa2021_assignment3.Rmd)) in Rstudio. 

- Right under each question, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas under Assignment 3. 

-------

Problems
-------
 \

1. Create the following data frame

| Rank 	| Age 	| Name  	|
|------	|-----	|-------	|
| 0    	| 28  	| Tom   	|
| 1    	| 34  	| Jack  	|
| 2    	| 29  	| Steve 	|
| 3    	| 42  	| Ricky 	|

```{r}
df <- data.frame(Rank=c(0,1,2,3), Age=c(28,34,29,42), Name=c('Tom','Jack','Steve','Ricky'))
df
```

2. Use `read.csv` to import the Covid19 Vaccination data from WHO: [link](https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv)

Show the names of the variables in the data

```{r}
df3 <- read.csv('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv')
names(df3)
```

3. How many columns and rows the data have? 

```{r}
length(df3$state) #for numbers of rows
length(df3) #for number of columns
dim(df3) # both in one function
```

4. How many missing values are there?  Show the missing values by columns.  What variable has the most number of missing values?

```{r}
sum(is.na(df3))
#there are no missing values
colSums(is.na(df3))
```

5. What is the class of the `date` column. Change the `date` columns to `date` type using the `as.Date` function. Show the new class of the `date` column. 

```{r}
class(df3$date)
df3$date <- as.Date(df3$date)
class(df3$date)
```

6. Capitalize the names of all the variables

```{r}
names(df3)[1] <- 'Date'
names(df3)[2] <- 'State'
names(df3)[3] <- 'Fips'
names(df3)[4] <- 'Cases'
names(df3)[5] <- 'Deaths'
```

7. Find the average number of cases per day. Find the maximum cases a day.  

```{r}
mean(df3$Cases)
max(df3$Cases)
```

8. How many states are there in the data?  
```{r}
length(unique(df3$State))
```


9. Create a new variable `weekdays` to store the weekday for each rows. 

```{r}
df3$weekdays <- weekdays(df3$Date)
str(df3)
```


10. Create the categorical variable `death2` variable taking the values as follows

- `has_death` if there is a death that day
-  `no_death` if there is no death that day

Find the frequency and relative frequency of `no_death` and `has_death`. 

```{r}
df3$death2 <- ifelse(df3$Deaths==0, 'no_death','has_death')
table(df3$death2)
mean(df3$death2=='has_death')
#96.93% of days have a death
```

11. Find the first quartile (Q1), second quartile (Q2) and and third quartile (Q3) of the variable `death`. (Hint:  Use the `summary` function)

```{r}
summary(df3$Deaths)
#Q1 = 364, Q2 = 2082, Q3 = 7380 as of 9/21
```


12. Create the categorical variable `death3` variable taking the values as follows

- `low_death` if the number of deaths smaller than the 25 percentile (Q1)

- `mid_death` if the number of deaths from Q1 to Q3

- `high_death` if the number of deaths greater than Q3

```{r}
df3$death3 <- cut(df3$Deaths, 
                      breaks = c(-Inf, 364, 7380,Inf), 
                      labels=c('low_death','mid_death','high_death'))
table(df3$death3)
```


13. Find the average cases in Rhode Island in 2021

```{r}
dfRI <- df3[df3$Date>='2021-01-01'&df3$State=='Rhode Island',]
#don't need to cap date yet because no 2022 data
by(dfRI$Cases, dfRI$State, mean)
```


14. Find the median cases by weekdays in Rhode Island in 2021 

```{r}
#dfRI already has data only in Rhode Island in 2021
by(dfRI$Cases,dfRI$weekdays, median)
```


15. Compare the median cases in Rhode Island in June, July, August and September in 2021. 

```{r}
dfRI$months <- months(dfRI$Date)
dfRI <- dfRI[dfRI$Date>='2021-06-01'&dfRI$Date<'2021-10-01',]
by(dfRI$Cases,dfRI$months, median)
```