
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 16: Writing a function - Loop"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](fa2021_assignment16.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

1. Write the following function. Give examples to test your function.

- Input: a data frame

- Output: a data frame with all the missing of numeric variables replaced by the associated means.

```{r, error=TRUE}
#This function works for an individual column regardless if it is numeric or categorical, but I could not correctly solve the loop portion

ReplaceBlanks <- function(d)
{
  library(ggplot2)
  for (i in 1:length(d))
  {
    if (is.numeric(d[[i]]))
        {print("Yes")
      library(tidyr)
        d[[i]] <- replace_na(d[[i]],mean(d[[i]], na.rm=TRUE))
        return(d[[i]])}
    else if (!is.numeric(d[[i]]))
    {print("No")
    mode_of_x <- names(sort(-table(d[[i]])))[1]
    library(tidyr)
    d[[i]] <- replace_na(d[[i]], mode_of_x) 
      return(d[[i]])}
    
  }
}

library(tidyverse)
dfTitanic <-read_csv("C:/Users/student/Desktop/Data Mining 460/titanic.csv")

sum(is.na(dfTitanic[[6]]))
sum(is.na(ReplaceBlanks(dfTitanic[[6]])))
sum(is.na(dfTitanic[[12]]))
sum(is.na(ReplaceBlanks(dfTitanic[[12]])))
#Works fine for one column at a time, won't go through all columns

colSums(is.na(dfTitanic))
colSums(is.na(ReplaceBlanks(dfTitanic)))
#Does not work because it is not a matrix, only one column apparently?

viewthisplease <-ReplaceBlanks(dfTitanic)
viewthisplease
#only views one column? not sure why the function only keeps the first column

```

Hint: [Similar function](https://bryantstats.github.io/math421/gganimate/18_writing_functions_fa21.html#28)

-------

2. Write the following function. Give examples to test your function.

- Input: a data frame

- Output: a data frame with all the missing of variables replaced by the associated means (for numeric variables) or modes (for non-numeric variables).

Hint: Use If-statement to combine the function in Problem 1 and the function in [this example](https://bryantstats.github.io/math421/gganimate/18_writing_functions_fa21.html#28)


```{r}
# Same answer as 1
```

-------

3. Write the following function. Give examples to test your function.

- Input: a data frame

- Output: Bar plots of all non-numeric variables

Hint: [Similar function](https://bryantstats.github.io/math421/gganimate/18_writing_functions_fa21.html#33)

```{r}
bar_plot_all <- function(d)
{
  library(ggplot2)
  for (i in 1:length(d))
  {
    if (!is.numeric(d[[i]]))
    {
      print(ggplot(d, aes(x = d[[i]]))+ 
              geom_bar()+
              labs(x = names(d)[i]))
    }
  }
}
library(tidyverse)
dfTitanic <-read_csv("C:/Users/student/Desktop/Data Mining 460/titanic.csv")

bar_plot_all(dfTitanic)
```



-------

4. Write the following function. Give examples to test your function.

- Input: a data frame

- Output: all possible the bar plots of a non-numeric variable filled by a non-numeric variable.

Hint: [Similar function](https://bryantstats.github.io/math421/gganimate/18_writing_functions_fa21.html#40)

```{r}
bar_plot_all2 <- function(d)
{
  library(ggplot2)
  l <- length(d)
  for (i in 1:(l-1))
    for (j in (i+1):l)
  {
    if (!is.numeric(d[[i]])& (!is.numeric(d[[j]])))
    {
      print(ggplot(d, aes(x = d[[i]], fill = d[[j]]))+ 
              geom_bar()+labs(x = names(d)[i], fill = names(d)[j]))
    }
  }
}
#This code works, but the resulting graphs for the Titanic dataset look ridiculous because there are many different categories that have very few values (like Name, Cabin, etc)
bar_plot_all2(dfTitanic)
```


-------

5. Write the following function. Give examples to test your function.

- Input: a data frame

- Output: 

  - all possible the bar plots of a non-numeric variable filled by a non-numeric variable.
  
  - all possible the density plots of a numeric variable colored by a non-numeric variable
  
  - all possible the scatter plots.

Hint: Combine [this function](https://bryantstats.github.io/math421/gganimate/18_writing_functions_fa21.html#37), [this function](https://bryantstats.github.io/math421/gganimate/18_writing_functions_fa21.html#40), and the function in Question 4. One way to combine is creating a new function, `quick_plot`, and call these three functions within `quic_kplot`. 

```{r, warning=FALSE}
AllThePlots <- function(d)
{
  library(ggplot2)
  l <- length(d)
  for (i in 1:(l-1))
    for (j in (i+1):l)
  {
    if (!is.numeric(d[[i]])& (!is.numeric(d[[j]])))
    {
      print(ggplot(d, aes(x = d[[i]], fill = d[[j]]))+ 
              geom_bar()+labs(x = names(d)[i], fill = names(d)[j]))
    }

else if (is.numeric(d[[i]])& (!is.numeric(d[[j]])))
    {
      print(ggplot(d, aes(x = d[[i]], color = d[[j]]))+ 
              geom_density()+labs(x = names(d)[i], color = names(d)[j]))
}
      else if (is.numeric(d[[i]])&is.numeric(d[[j]]))
    {
      print(ggplot(d, aes(x = d[[i]], y = d[[j]]))+ 
              geom_point()+
              labs(x = names(d)[i], y = names(d)[j]))
    }
      
    }
}
dfTitanicSmall <- select(dfTitanic,-Name,-PassengerId,-SibSp,-Parch)
#Just to remove some columns
AllThePlots(dfTitanic)
```

