
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 9: Visualization - Revealing a graph"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](fa2021_assignment9.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


-------

1. Use the dataset by of covid 19 by WHO at https://covid19.who.int/WHO-COVID-19-global-data.csv. Find the three countries with the most numbers of deaths by Covid-19. 

Hint: 

-  Use `filter` to select the most current date 

-  Pipe with `arrange(-Cumulative_deaths)` to rank the data by the number of cumulative deaths

```{r}
library(tidyverse)
library(gganimate)
df<- read_csv('https://covid19.who.int/WHO-COVID-19-global-data.csv')

df %>% filter(Date_reported=='2021-10-18') %>% arrange(-Cumulative_deaths)
#Top three most deaths is India, Russian Federation, and Indonesia
```


2. Make a plot revealing the number of deaths in the three countries with the most numbers of deaths
```{r}
df %>% 
  filter(Country %in% c('India','Russian Federation','Indonesia')) %>% 
  ggplot(aes(y=Cumulative_deaths,
            x=Date_reported, color=Country))+ 
  geom_line()+
  transition_reveal(Date_reported)
```


3. Create the new variable (column) `death_per_cases` recording the number of deaths per cases (Hint: divide cumulative deaths by cumulative cases).  What are the three countries with the highest deaths per cases? 

```{r}
df$death_per_cases <- df$Cumulative_deaths/df$Cumulative_cases
df %>% filter(Date_reported=='2021-10-18') %>% arrange(-death_per_cases)
#Top three worst deaths per cases (as of Oct 18, 2021) are Yemen, Sudan, and Syrian Arab Republic
```

4. Make a plot revealing the number of deaths per cases of the US, Italy and Mexico. 
```{r}
df$dpcleft = substr(df$death_per_cases,1,6)
# makes the number labels look cleaner
df %>% 
  filter(Country %in% c('Italy','Mexico','United States of America')) %>% 
  ggplot(aes(y=death_per_cases,
            x=Date_reported, color=Country))+ 
  geom_line() +
  geom_text(aes(label = dpcleft), 
            hjust = -.1, size=5) +
  transition_reveal(Date_reported)
```


5. Use `transition_reveal` to create three animated graph with the vaccination data at 

https://raw.githubusercontent.com/govex/COVID-19/master/data_tables/vaccine_data/us_data/time_series/vaccine_data_us_timeline.csv

```{r}
df5 <-read_csv('https://raw.githubusercontent.com/govex/COVID-19/master/data_tables/vaccine_data/us_data/time_series/vaccine_data_us_timeline.csv')
```

```{r}
df5 %>% filter(Vaccine_Type =='All') %>% 
  filter(Province_State %in% c('Rhode Island', 'Massachusetts', 'Vermont','Maine','Connecticut','New Hampshire')) %>% 
  ggplot(aes(y=Stage_Two_Doses,
            x=Date, color=Province_State))+ 
  geom_line()+
  transition_reveal(Date) + labs(x='Date', 
       y = 'Cumulative Doses Administered', 
       color='State', 
       title = 'New England Vaccines', 
       caption = 'Massachusetts administered many more vaccines.')
```

```{r}
df5 %>% filter(Vaccine_Type != 'All') %>% 
  filter(Province_State %in% c('Rhode Island')) %>% 
  ggplot(aes(y=Doses_admin,
            x=Date, color=Vaccine_Type))+ 
  geom_line()+
  transition_reveal(Date) + labs(x='Date', 
       y = 'Doses Administered', 
       color='Vaccine Brand', 
       title = 'Rhode Island Vaccine Types', 
       caption = 'More Pfizer vaccines were administered.')
```

```{r}
df5 %>% filter(Date=='2021-10-17') %>% filter(Vaccine_Type == 'All') %>% arrange(-Stage_One_Doses)
#Top five most vaccines (as of Oct 17, 2021) are California, Texas, Florida, New York, Pennsylvania

df5 %>% filter(Vaccine_Type == 'All') %>% 
  filter(Province_State %in% c('California', 'Texas','Florida','New York','Pennsylvania')) %>% 
  ggplot(aes(y=Stage_One_Doses,
            x=Date, color=Province_State))+ 
  geom_line()+
  transition_reveal(Date) + labs(x='Date', 
       y = 'First Doses Administered', 
       color='State', 
       title = 'First Doses in Large States', 
       caption = 'Florida and New York lines look similar.')
```

