
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "How effective the Covid vaccines?"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](fa2021_assignment9_extra_credits.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


-------

How effective is the covid vaccines? There are multiple analysis to answer the question. This following practice give ONE way to analyze how effective the Covid vaccines are. The analysis is done for the United States. 

1. Import the two following data

  - Data 1: https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations.csv

  - Data 2: https://covid19.who.int/who-data/vaccination-data.csv
```{r}
library(tidyverse)
library(gganimate)
df1 <- read_csv('https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations.csv')
df2 <- read_csv('https://covid19.who.int/who-data/vaccination-data.csv')
df3 <-df<- read_csv('https://covid19.who.int/WHO-COVID-19-global-data.csv')
#Data set 2 contains no information about deaths, so I am using this dataset (3) instead.  It also includes the column 'Date_reported' used in step 2.
```


2. We will merge these two datasets in one for our analysis.  We will use the `inner_join()` function to join the two datasets into one new datasets where the primary keys are Country and Date.  First we need to change the names of the two datasets so that they have the same name for country and date.  In dataset 1, change the name of the `location` to `Country`. In dataset 2, change the name of the variable `Date_reported` to `date`
```{r}
df1$Country <-df1$location
df2$date <- df2$DATE_UPDATED
df3$date <-df3$Date_reported
```


3. In dataset 2 and variable `Country` change `United States of America` to `United States` (You can use `str_replace`)
```{r}
df3$Country <- str_replace(df3$Country,'United States of America','United States')
```


4. Use `inner_join(data1, data2, by = c('Country','date')` to create a new dataset. 
```{r}
df99 <- inner_join(df1, df3, by = c('Country','date'))
```


For more about joining data, see this datacamp course:  https://app.datacamp.com/learn/courses/joining-data-with-dplyr-in-r

5. In the new dataset, create a variable `death_per_vac` by dividing the `Cumulative_deaths` by `total_vaccinations`. 

```{r}
df99$death_per_vac = df99$Cumulative_deaths/df99$total_vaccinations
```

6. Plot the `death_per_vac` by `date` for the United State using `geom_point`. 
```{r}
df99 %>% 
  filter(Country == 'United States') %>% 
  ggplot(aes(y=death_per_vac,
            x=date))+ 
  geom_point()
#Fairly low amount of deaths compared to vaccines
```


7. Use the log transformation for the plot above for better view.  Hint: add to the code above `+scale_y_continuous(trans='log')`
```{r}
df99 %>% 
  filter(Country == 'United States') %>% 
  ggplot(aes(y=death_per_vac,
            x=date))+ 
  geom_point() + scale_y_continuous(trans='log')
```


8. Filter the plot above for July to current date only. Give your comments.

```{r}
df99 %>% 
  filter(Country == 'United States') %>% filter(date > '2021-06-30') %>% 
  ggplot(aes(y=death_per_vac,
            x=date))+ 
  geom_point() + scale_y_continuous(trans='log')
#Deaths per vaccine is low, but seems like may be increasing again as the weather gets colder
```

9. Do the above analysis to other countries or continents. Give your comments.
```{r}
df99 %>% group_by(WHO_region, date) %>% summarise(DPV=mean(death_per_vac)) %>% 
  ggplot(aes(y=DPV,
            x=date, color=WHO_region))+ 
  geom_point() + scale_y_continuous(trans='log')
#Very many more deaths per vaccines in other regions of the world than Americas at start.  All regions seem to have lower death:vaccine ratios as time progresses.
```
```{r}
df99 %>% filter(WHO_region=='SEARO') %>% ggplot(aes(y=death_per_vac,
            x=date, color=Country))+ 
  geom_point() + scale_y_continuous(trans='log')
#All countries in Southeast Asia seem to have decreasing deaths per vaccines overtime too
```


10. Do you have any other analysis for the effectiveness of the vaccine covid?
```{r}
df99$CasePerVac = df99$New_cases / df99$people_vaccinated

df99 %>% filter(Country %in% c('India','United States', 'Italy','Mongolia')) %>% filter(date>'2021-01-31') %>% ggplot(aes(y=CasePerVac,
            x=date, color=Country))+ 
  geom_point(shape=16)
#Over time there are less total cases per vaccine in different countries throughout the world.  The ratio seems to mostly stabilize over the Summer, but cases in the US may be increasing now that it is almost Winter again.
```

