---
title: "Final Shiny Apps"
author: "Matthew Melnychuk"
date: "12/6/2021"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1: Inverse Norm

```{r}
library(shiny)

######################################
# Set User Interface
ui <- fluidPage(
  titlePanel("Hello Shiny!"),
  
  sidebarLayout(
    
    sidebarPanel(
      sliderInput(
        inputId = "var1",
        label = "Decide a number",
        min = 1,
        max = 100,
        value=10
      )
      
    ),
    
    mainPanel(
      # Output: Histogram ----
      plotOutput(outputId = 'blah1')
    )
    
  )
  
)


######################################
# Main codes for the app
server <- function(input, output) {
  
  output$blah1 <-renderPlot({
    m = input$var1
    hist(rnorm(mean=m, n=1000))
  }
  )
  
  
}

######################################
# Run the app
shinyApp(ui = ui, server = server)

```


## Question 2: Titanic Barplot

```{r}
library(tidyverse)
library(shiny)

d = read_csv('titanic.csv')
# drop some columns
d <- d %>% select(-Name, -PassengerId, -Ticket)

# convert categorical variables to character type
d <- d %>% mutate_at(c('Survived', 'Pclass'), as.character)

# get variable names
variables_names = names(d)

# get names of numeric variables
variables_numeric = d %>% select_if(is.numeric) %>% names

# get names of categorical variables
variables_category = d %>% select_if(is.character) %>% names

ui <- fluidPage(
  
  titlePanel("Barplot for Titanic Data"),
  
  sidebarLayout(
    
    sidebarPanel(
      
      selectInput(
        inputId ="var1",
        label = "Select a Categorical Variable",
        choices = variables_category, selected = "Age"
      ),
      
      selectInput(
        inputId ="var2",
        label = "Select a Categorical Variable",
        choices = variables_category,
        selected = "Sex"
      )
      
    ),
    
    # Main panel for displaying outputs ----
    mainPanel(
      
      # Output: Histogram ----
      
      plotOutput(outputId = 'show_plot')
    )
  )
)

# server is a function! 
server <- function(input, output) {
  
  
  output$show_plot <- renderPlot({
    d = read_csv('titanic.csv')
    v1 = input$var1
    v2 = input$var2
   
    
    library(ggplot2)
      
    r = ggplot(d, aes(x = d[[v1]], fill = as.factor(d[[v2]])))+
        geom_bar()+
        labs(x = v1, fill = v2)
      
    return(r)
    
    
  })
  
}
# app
shinyApp(ui = ui, server = server)


```

## Question 3: Census Density with Checkbox

```{r}
library(tidyverse)
library(shiny)

dd = read_csv('adult_modified.csv')

# get variable names
variables_names = names(dd)

# get names of numeric variables
variables_numeric = dd %>% select_if(is.numeric) %>% names

# get names of categorical variables
variables_category = dd %>% select_if(is.character) %>% names



ui <- fluidPage(
  
  titlePanel("Density Plot for Census Data"),
  
  sidebarLayout(
    
    sidebarPanel(
      
      selectInput(
        inputId ="var1",
        label = "Select a Numeric Variable",
        choices = variables_numeric, selected = "age"
      ),
      
      selectInput(
        inputId ="var2",
        label = "Select a Categorical Variable",
        choices = variables_category,
        selected = "sex"
      ),
      
      checkboxGroupInput(inputId = "native__country", label = "Select Native Country",
                         choices = names(table(dd$native_country)), selected="US", inline = TRUE),
      
    ),
    
    # Main panel for displaying outputs ----
    mainPanel(
      
      # Output: Histogram ----
      
      plotOutput(outputId = 'show_plot')
    )
  )
)

# server is a function! 
server <- function(input, output) {
  
  
  output$show_plot <- renderPlot({
    
    v1 = input$var1
    v2 = input$var2
   
    dd <- dd %>% filter(native_country %in% input$native__country)
    
    library(ggplot2)
    
      ggplot(dd, aes(x = dd[[v1]], color = as.factor(dd[[v2]])))+
        geom_density()+
        labs(x = v1, color = v2)
    
    
  })
  
}
# app
shinyApp(ui = ui, server = server)
```


## Question 4: Census Density with Slider Input
```{r}
library(tidyverse)
library(shiny)

dd = read_csv('adult_modified.csv')


# get variable names
variables_names = names(dd)

# get names of numeric variables
variables_numeric = dd %>% select_if(is.numeric) %>% names

# get names of categorical variables
variables_category = dd %>% select_if(is.character) %>% names


ui <- fluidPage(
  
  titlePanel("Density Plot for Census Data"),
  
  sidebarLayout(
    
    sidebarPanel(
      
      selectInput(
        inputId ="var1",
        label = "Select a Numeric Variable",
        choices = variables_numeric, selected = "age"
      ),
      
      selectInput(
        inputId ="var2",
        label = "Select a Categorical Variable",
        choices = variables_category,
        selected = "sex"
      ), 
      
      sliderInput(inputId = "Hours",
                  "Select Hours per Week:",
                  min = min(dd$hours_per_week, na.rm=TRUE),
                  max = max(dd$hours_per_week, na.rm=TRUE),
                  value= c(20, 45))
      
    ),
    
    # Main panel for displaying outputs ----
    mainPanel(
      
      # Output: Histogram ----
      
      plotOutput(outputId = 'show_plot')
    )
  )
)

# server is a function! 
server <- function(input, output) {
  
  
  output$show_plot <- renderPlot({
    
    v1 = input$var1
    v2 = input$var2
   
    
    library(ggplot2)
    
    dd <- dd %>% filter(hours_per_week>input$Hours[1], hours_per_week<input$Hours[2])
      
    ggplot(dd, aes(x = dd[[v1]], color = as.factor(dd[[v2]])))+
        geom_density()+
        labs(x = v1, color = v2)
      
    
  })
  
}
# app
shinyApp(ui = ui, server = server)
```


## Question 5: Covid Date Range Input

```{r}
library(tidyverse)
library(shiny)

d = read_csv('who_covid.csv')
numeric_variables = d %>% select_if(is.numeric) %>% names()

variables_names = names(d)

ui <- fluidPage(
  
  titlePanel("Covid19 Data by Countries"),
  
  sidebarLayout(
    
    # Side Panel for reading inputs
    sidebarPanel(
      
      selectInput(
        inputId ="var1",
        label = "Select a statistic",
        choices = numeric_variables, selected = "Cumulative_cases"
      ),
      
      selectInput(inputId = "selected_country", label = "Select Country",
                         choices = names(table(d$Country)), selected = "Spain"), 
      
      selectInput(inputId = "selected_country2", label = "Select Country",
                  choices = names(table(d$Country)), selected = "Italy"), 
      
      selectInput(inputId = "selected_country3", label = "Select Country",
                  choices = names(table(d$Country)), selected = "Germany"),
      
      
      dateRangeInput(inputId = "date", 
                     strong("Date range"), 
                     start = "2020-01-01", end = "2021-11-25",
                     min = "2020-01-01", max = "2021-11-25"),
    ),
    
    # Main panel for displaying outputs ----
    mainPanel(
      plotOutput(outputId = 'show_plot')
    )
  )
)

# server is a function! 
server <- function(input, output) {
  
  output$show_plot <- renderPlot({
    d = read_csv('who_covid.csv')
    v1 = input$var1
    v2 = input$var2
    country1 = input$selected_country
    country2 = input$selected_country2
    country3 = input$selected_country3
    
    country = c(country1,country2,country3)
    
    print(v1)
    library(ggplot2)
    # r <- d %>% filter(Country %in% country) %>% 
    #   ggplot(aes(x = Date_reported, y = d[[v1]]))+
    #   geom_point()+
    #   labs(x = 'Date', y = d[[v1]])
    
    d <- d %>% filter(Country %in% country, Date_reported>input$date[1],  Date_reported<input$date[2]) 
    
    r <- d %>% ggplot(aes(x = Date_reported, y = d[[v1]], color = Country))+
      geom_point()+
      labs(x = 'Date', y = v1)
    
    return(r)
    
    
  })
  
}
# app
shinyApp(ui = ui, server = server)
```

## Question 6: Census Plots with Radio Buttons

```{r}
library(tidyverse)
library(shiny)

dd = read_csv('adult_modified.csv')

# get variable names
variables_names = names(dd)

# get names of numeric variables
variables_numeric = dd %>% select_if(is.numeric) %>% names

# get names of categorical variables
variables_category = dd %>% select_if(is.character) %>% names


ui <- fluidPage(
  
  titlePanel("Density Plot for Census Data"),
  
  sidebarLayout(
    
    sidebarPanel(
      
      selectInput(
        inputId ="var1",
        label = "Select a Numeric Variables",
        choices = variables_numeric, selected = "age"
      ),
      
      selectInput(
        inputId ="var2",
        label = "Select a Categorical Variables",
        choices = variables_category,
        selected = "sex"
      ),
      
      radioButtons(inputId = "plot_choice", 
                   label = h3("Select Plot:"),
                   choices = c("Density Plot" = "density",
                     "Histogram Plot" = "histogram"),
                   selected = 'density')
      
    ),
    
    # Main panel for displaying outputs ----
    mainPanel(
      
      # Output: Histogram ----
      
      plotOutput(outputId = 'show_plot')
    )
  )
)

# server is a function! 
server <- function(input, output) {
  
  
  output$show_plot <- renderPlot({
    
    v1 = input$var1
    v2 = input$var2
   
    
    library(ggplot2)
    
    if(input$plot_choice == 'density')
      
    {
      ggplot(dd, aes(x = dd[[v1]], color = as.factor(dd[[v2]])))+
        geom_density()+
        labs(x = v1, color = v2)
    }
    
    else
    {
      ggplot(dd, aes(x = dd[[v1]], color = as.factor(dd[[v2]])))+
        geom_histogram()+
        labs(x = v1, color = v2)
    }
      
      
    
  })
  
}
# app
shinyApp(ui = ui, server = server)
```

## Question 7: Census Plots with NavBar
```{r}
library(tidyverse)
library(shiny)

dd = read_csv('adult_modified.csv')

# get variable names
variables_names = names(dd)

# get names of numeric variables
variables_numeric = dd %>% select_if(is.numeric) %>% names

# get names of categorical variables
variables_category = dd %>% select_if(is.character) %>% names

ui <- navbarPage("Navigation",
           tabPanel("Numeric Density",
                    
                    sidebarLayout(
                      sidebarPanel(
                        
                        selectInput(
                          inputId ="var1",
                          label = "Select a Numeric Variable",
                          choices = variables_numeric, selected = "age"
                        )
                      ),
                      
                      mainPanel(
                        plotOutput(outputId = 'show_plot')
                      )
                    )
           ),
           
           
           tabPanel("Categorical Count",
                    
                    sidebarLayout(
                      sidebarPanel(
                        
                        selectInput(
                          inputId ="var2",
                          label = "Select a Categorical Variable",
                          choices = variables_category, selected = "sex"
                        )
                      ),
                      
                      mainPanel(
                        plotOutput(outputId = 'show_plot2')
                      )
                    )
           )
)

# server is a function! 
server <- function(input, output) {
  
  
  output$show_plot <- renderPlot({
    
    v1 = input$var1
    
    library(ggplot2)
    
      ggplot(dd, aes(x = dd[[v1]]))+
        geom_density()+
        labs(x = v1)
    
    
  })
  
  output$show_plot2 <- renderPlot({
    
    v2 = input$var2
    
    library(ggplot2)
    
    ggplot(dd, aes(x = dd[[v2]]))+
      geom_bar()+
      labs(x = v2)
    
    
  })
  
}
# app
shinyApp(ui = ui, server = server)
```

## Question 8: Bar Plots for Uploaded Data
```{r}
library(tidyverse)
library(shiny)
library(DT)

ui <- fluidPage(
  
  titlePanel("Bar Plots for Uploaded Data"),
  
  sidebarLayout(
    
    sidebarPanel(
      fileInput('f1', label = 'Upload data for visualization', accept = '.csv'),
      
      selectInput('v1', label='Select a Categorical Variable', choices=''),
      selectInput('v2', label='Select a Categorical Variable', choices='')
    
      
    ),
    
    # Main panel for displaying outputs ----
    mainPanel(
      plotOutput(outputId = 'show_plot')
    )
  )
)

# server
server <- function(input, output, session) {
  
  myData <- reactive({
    inFile = input$f1
    if (is.null(inFile)) return(NULL)
    data <- read.csv(inFile$datapath, header = TRUE)
    data
  }
    )
    
  output$show_plot <- renderPlot({
  
    inFile = input$f1
    if (!is.null(inFile))
    {
    v1 = input$v1
    d <- read.csv(inFile$datapath, header = TRUE)
    
    v1 = input$v1
    v2 = input$v2
    
    
      library(ggplot2)
      
      r = ggplot(d, aes(x = d[[v1]], fill = as.factor(d[[v2]])))+
        geom_bar()+
        labs(x = v1, fill = v2)
      
    return(r)
    }
    else{
      return(NULL)
    }
  })
  
  
  observeEvent(input$f1,{ 
    inFile = input$f1
    data <- read.csv(inFile$datapath, header = TRUE)   
               updateSelectInput(session, 'v1', choices = names(data))}
               )
  
  observeEvent(input$f1,{ 
    inFile = input$f1
    data <- read.csv(inFile$datapath, header = TRUE)   
    updateSelectInput(session, 'v2', choices = names(data))}
  )

}


shinyApp(ui = ui, server = server)
```


## Question 9: Scatter Plots for Uploaded Data

```{r}
library(tidyverse)
library(shiny)
library(DT)

ui <- fluidPage(
  
  titlePanel("Scatter Plots for Uploaded Data"),
  
  sidebarLayout(
    
    sidebarPanel(
      fileInput('f1', label = 'Upload data for visualization', accept = '.csv'),
      
      selectInput('v1', label='Select a Numerical Variable for X', choices=''),
      selectInput('v2', label='Select a Numerical Variable for Y', choices=''),
      selectInput('v3', label='Select a Categorical Variable', choices='')
      
    ),
    
    # Main panel for displaying outputs ----
    mainPanel(
      plotOutput(outputId = 'show_plot')
    )
  )
)

# server
server <- function(input, output, session) {
  
  myData <- reactive({
    inFile = input$f1
    if (is.null(inFile)) return(NULL)
    data <- read.csv(inFile$datapath, header = TRUE)
    data
  }
    )
    
  output$show_plot <- renderPlot({
  
    inFile = input$f1
    if (!is.null(inFile))
    {
    v1 = input$v1
    d <- read.csv(inFile$datapath, header = TRUE)
    
    v1 = input$v1
    v2 = input$v2
    v3 = input$v3
    
    
      library(ggplot2)
      
      r = ggplot(d, aes(x = d[[v1]], y = d[[v2]],color = d[[v3]]))+
        geom_point()+
        labs(x = v1, y = v2, color = v3)
      
    return(r)}
    
    else{
      return(NULL)
    }
    
  })
  
  
  observeEvent(input$f1,{ 
    inFile = input$f1
    data <- read.csv(inFile$datapath, header = TRUE)   
               updateSelectInput(session, 'v1', choices = names(data))}
               )
  
  observeEvent(input$f1,{ 
    inFile = input$f1
    data <- read.csv(inFile$datapath, header = TRUE)   
    updateSelectInput(session, 'v2', choices = names(data))}
  )

  observeEvent(input$f1,{ 
    inFile = input$f1
    data <- read.csv(inFile$datapath, header = TRUE)   
    updateSelectInput(session, 'v3', choices = names(data))}
  )
}


shinyApp(ui = ui, server = server)
```


## Question 10: NCAA Volleyball Stats Scatter Plots
```{r}
library(tidyverse)
library(shiny)

dvb = read_csv('ncaaClean.csv')

# get variable names
variables_names = names(dvb)

# get names of numeric variables
variables_numeric = dvb %>% select_if(is.numeric) %>% names

# get names of categorical variables
variables_category = dvb %>% select_if(is.character) %>% names


ui <- fluidPage(
  
  titlePanel("Plots for NCAA Volleyball Statistics 2019"),
  
  sidebarLayout(
    
    sidebarPanel(
      
      selectInput(
        inputId ="var1",
        label = "Select a Numeric Variable (for X)",
        choices = variables_numeric, selected = "HitPercent"
      ),
      
      selectInput(
        inputId ="var2",
        label = "Select a Numeric Variable (for Y)",
        choices = variables_numeric,
        selected = "Blocks"
      ), 
      
      radioButtons(inputId = "plot_choice", 
                   label = h3("Select Plot:"),
                   choices = c("Scatter Plot" = "scatter",
                               "Density Plot (only uses 1st var)" = "density"),
                   selected = 'scatter'),
      
      checkboxGroupInput(inputId = "selected_region", label = "Region",
                         choices = c('East','Midwest','South','Southeast','West'), selected = c("East","South"), inline = TRUE),
      
      sliderInput(inputId = "winP",
                  "Select Win % of Interest:",
                  min = min(dvb$WinPercent, na.rm=TRUE),
                  max = max(dvb$WinPercent, na.rm=TRUE),
                  value= c(0.25, 0.75))
      
    ),
    
    # Main panel for displaying outputs ----
    mainPanel(
      
      # Output: Histogram ----
      
      plotOutput(outputId = 'show_plot')
    )
  )
)

# server is a function! 
server <- function(input, output) {
  
  
  output$show_plot <- renderPlot({
    
    v1 = input$var1
    v2 = input$var2
    region = input$selected_region
   
    
    library(ggplot2)
    
    dvb <- dvb %>% filter(Region %in% region,WinPercent>input$winP[1], WinPercent<input$winP[2])
    
    if(input$plot_choice == 'scatter')
      
    {
      ggplot(dvb, aes(x = dvb[[v1]], y = dvb[[v2]], color = as.factor(dvb$Conference)))+
        geom_point()+
        labs(x = v1, y = v2, color = 'Conference')
    }
    
    else
    {
      r <- dvb %>% ggplot(aes(x = dvb[[v1]], color = dvb$Conference))+
        geom_density()+
        labs(x = v1, color = 'Conference')
      return(r)
    }
    
    
    
  })
  
}
# app
shinyApp(ui = ui, server = server)
```


## Question 11: NEC Volleyball Standings Over Time

```{r}
library(tidyverse)
library(shiny)

d = read_csv('NEC_Standings2.csv')
numeric_variables = d %>% select_if(is.numeric) %>% names()
d$SeasonDate <- as.Date(d$SeasonDate,"%m/%d/%y")

variables_names = names(d)

ui <- fluidPage(
  
  titlePanel("Northeast Conference Volleyball"),
  
  sidebarLayout(
    
    # Side Panel for reading inputs
    sidebarPanel(
      
      selectInput(
        inputId ="var1",
        label = "Select a statistic",
        choices = numeric_variables, selected = "Conf Win"
      ),
      
      radioButtons(inputId = "plot_choice", 
                   label = h3("Select Plot:"),
                   choices = c("Scatter Plot" = "scatter",
                               "Line Plot" = "line"),
                   selected = 'scatter'),
      
      checkboxGroupInput(inputId = "selected_playoff", label = "Playoff Team?",
                         choices = c('Yes','No'), selected = "Yes", inline = TRUE),
      
      dateRangeInput(inputId = "Date", 
                     strong("Date range"), 
                     start = "2017-01-01", end = "2021-12-31",
                     min = "1997-01-01", max = "2021-12-31"),
    ),
    
    # Main panel for displaying outputs ----
    mainPanel(
      plotOutput(outputId = 'show_plot')
    )
  )
)

# server is a function! 
server <- function(input, output) {
  
  output$show_plot <- renderPlot({
    d = read_csv('NEC_Standings2.csv')
    v1 = input$var1
    v2 = input$var2
    playoff = input$selected_playoff
    print(v1)
    library(ggplot2)
    
    d <- d %>% filter(Playoffs %in% playoff, SeasonDate>input$Date[1],  SeasonDate<input$Date[2]) 
    
    
    if(input$plot_choice == 'scatter')
      
    {
      library(lubridate)
      r <- d %>% ggplot(aes(x = year(SeasonDate), y = d[[v1]], color = School))+
        geom_point()+
        labs(x = 'Date', y = v1)
      return(r)
    }
    
    else
    {
      library(lubridate)
      r <- d %>% ggplot(aes(x = year(SeasonDate), y = d[[v1]], color = School))+
        geom_line()+geom_point()+
        labs(x = 'Date', y = v1)
      return(r)
    }
    
    
    
    
  })
  
}
# app
shinyApp(ui = ui, server = server)
```


## Question 12: Many Volleyball Plots Using NavBar
```{r}
library(tidyverse)
library(shiny)

dvb = read_csv('ncaaClean.csv')

dvb$FiveHun <- ifelse(dvb$WinPercent>.5, 'Over500','No')

# convert categorical variables to character type
dvb <- dvb %>% mutate_at(c('Tourney2018', 'PrevChamp','MenProgram'), as.character)

# get variable names
variables_names = names(dvb)

# get names of numeric variables
variables_numeric = dvb %>% select_if(is.numeric) %>% names

# get names of categorical variables
variables_category = dvb %>% select_if(is.character) %>% names

ui <- navbarPage("VB Plot Navigation",
                 tabPanel("Stat Density Plot",
                          
                          sidebarLayout(
                            sidebarPanel(
                              
                              selectInput(
                                inputId ="var1",
                                label = "Select a Numeric Variable",
                                choices = variables_numeric, selected = "Digs"
                              ),
                              
                              checkboxGroupInput(inputId = "playlast", label = "1 = Team Made Playoffs Last Year, 0 = No Playoffs Last Year",
                                                 choices = names(table(dvb$Tourney2018)), selected="1", inline = TRUE),
                              
                              radioButtons(inputId = "team500", 
                                           label = h3("Select Teams:"),
                                           choices = c("All Teams" = "allteam",
                                                       "Teams Over 0.500 Win%" = "winners"),
                                           selected = 'allteam'),
                              
                              sliderInput(inputId = "coachtenure",
                                          "Select length of Coach Tenure:",
                                          min = min(dvb$CoachTenure),
                                          max = max(dvb$CoachTenure),
                                          value= c(3, 12))
                            ),
                            
                            
                            mainPanel(
                              plotOutput(outputId = 'show_plot')
                            )
                          )
                 ),
                 
                 
                 tabPanel("Categorical Count",
                          
                          sidebarLayout(
                            sidebarPanel(
                              
                              selectInput(
                                inputId ="var2",
                                label = "Select a Categorical Variable",
                                choices = variables_category, selected = "Region"
                              ),
                              
                              checkboxGroupInput(inputId = "playlast2", label = "1 = Team Made Playoffs Last Year, 0 = No Playoffs Last Year",
                                                 choices = names(table(dvb$Tourney2018)), selected="1", inline = TRUE),
                              radioButtons(inputId = "team5002", 
                                           label = h3("Select Teams:"),
                                           choices = c("All Teams" = "allteam",
                                                       "Teams Over 0.500 Win%" = "winners"),
                                           selected = 'allteam'),
                              
                              sliderInput(inputId = "coachtenure2",
                                          "Select length of Coach Tenure:",
                                          min = min(dvb$CoachTenure),
                                          max = max(dvb$CoachTenure),
                                          value= c(3, 12))
                            ),
                            
                            
                            mainPanel(
                              plotOutput(outputId = 'show_plot2')
                            )
                          )
                 ),
                 
                 tabPanel("Win Percentage by X Statistic",
                          
                          sidebarLayout(
                            sidebarPanel(
                              
                              selectInput(
                                inputId ="var3",
                                label = "Select a Numeric Variable",
                                choices = variables_numeric, selected = "OppHitPercent"
                              ),
                              
                              checkboxGroupInput(inputId = "playlast3", label = "1 = Team Made Playoffs Last Year, 0 = No Playoffs Last Year",
                                                 choices = names(table(dvb$Tourney2018)), selected="1", inline = TRUE),
                              
                              radioButtons(inputId = "team5003", 
                                           label = h3("Select Teams:"),
                                           choices = c("All Teams" = "allteam",
                                                       "Teams Over 0.500 Win%" = "winners"),
                                           selected = 'allteam'),
                              
                              sliderInput(inputId = "coachtenure3",
                                          "Select length of Coach Tenure:",
                                          min = min(dvb$CoachTenure),
                                          max = max(dvb$CoachTenure),
                                          value= c(3, 12))
                            ),
                            
                            mainPanel(
                              plotOutput(outputId = 'show_plot3')
                            )
                          )
                 )
)

# server is a function! 
server <- function(input, output) {
  
  
  output$show_plot <- renderPlot({
    
    v1 = input$var1
    
    dvb <- dvb %>% filter(CoachTenure>input$coachtenure[1], CoachTenure<input$coachtenure[2])
    dvb <- dvb %>% filter(Tourney2018 %in% input$playlast)
    
    if(input$team500 == 'winners')
      
    {
      dvb <- dvb %>% filter(FiveHun == 'Over500')
    }
    
    else
    {
    }
    
    
    
    library(ggplot2)
    
    ggplot(dvb, aes(x = dvb[[v1]]))+
      geom_density()+
      labs(x = v1, y = 'Density')
    
    
  })
  
  output$show_plot2 <- renderPlot({
    
    v2 = input$var2
    
    dvb <- dvb %>% filter(CoachTenure>input$coachtenure2[1], CoachTenure<input$coachtenure2[2])
    dvb <- dvb %>% filter(Tourney2018 %in% input$playlast2)
    
    if(input$team5002 == 'winners')
      
    {
      dvb <- dvb %>% filter(FiveHun == 'Over500')
    }
    
    else
    {
    }
    
    
    library(ggplot2)
    
    ggplot(dvb, aes(x = dvb[[v2]]))+
      geom_bar()+
      labs(x = v2, y = 'Frequency')
    
    
  })
  
  output$show_plot3 <- renderPlot({
    
    v3 = input$var3
    
    dvb <- dvb %>% filter(CoachTenure>input$coachtenure3[1], CoachTenure<input$coachtenure3[2])
    dvb <- dvb %>% filter(Tourney2018 %in% input$playlast3)
    
    if(input$team5003 == 'winners')
      
    {
      dvb <- dvb %>% filter(FiveHun == 'Over500')
    }
    
    else
    {
    }
    
    
    library(ggplot2)
    
    ggplot(dvb, aes(x = dvb[[v3]], y = dvb$WinPercent))+
      geom_point()+
      labs(x = v3, y = 'Win Percentage')
    
    
  })
  
}
# app
shinyApp(ui = ui, server = server)
```
